Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-12-13 article.qmd
As TODO and bold to pending things

To overcome these limitations,
we developed poincaré and SimBio,
open-source Python packages for defining and simulating systems.
Poincaré allows defining differential equation systems,
while SimBio builds on it for defining reaction networks.
They are focused on providing an ergonomic experience to end-users by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
Since they are coded in Python,
every part from model definition to simulation can be extended from Python syntax.
Being the first-ever pure Python packages for systems modelling,
they offer extensive extensibility,
from simple tasks like reusing integrators defined in other packages,
to complex ones like altering the compilation process to leverage some structure in the equations (For example...).
The models built using these packages can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Furthermore,
they have a modular architecture with a clear separation of concerns,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
We showcased the reliability of these tools by benchmarking them against the simulation results from other tools.
We also highlighted the substantial performance improvements our tools offer,
as this is crucial for construction and simulation of models of whole cells and organisms,
which necessitate the simulation of significantly large-scale models.

@@

For the reaction networks,
our focus is on first-order differential equations
that describe the rate of change of species.
SimBio simplifies the definition of these network models by introducing `Species`,
which consists of a `poincare.Variable` and stoichiometry,
and `Reaction`,
a construct that converts reactants into products taking into account the stoichiometry (@fig-simbio).
Additionally, SimBio features `MassAction`, a subclass of `Reaction`,
which intuitively incorporates reactants into the rate law (@fig-simbio).
(is `RateLaw` not better term than `MassAction` ?)

@@

`Gen Comment: I amnto sure whether the number of figures are already over the limit, you may have to combine figures, while doing so, if space permits create a small schematics / summary figure, with icons and texts showing the significance / and applicability of the tools developed.`

>>>>

To overcome these limitations,
we developed poincaré and SimBio,
open-source Python packages for defining and simulating systems.
Poincaré allows defining differential equation systems,
while SimBio builds on it for defining reaction networks.
They are focused on providing an ergonomic experience to end-users by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
Since they are coded in Python,
every part from model definition to simulation can be extended from Python syntax.
Being the first-ever pure Python packages for systems modelling,
they offer extensive extensibility,
from simple tasks like reusing integrators defined in other packages,
to complex ones like altering the compilation process to leverage some structure in the equations.
**TODO: For example...**
The models built using these packages can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Furthermore,
they have a modular architecture with a clear separation of concerns,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
We showcased the reliability of these tools by benchmarking them against the simulation results from other tools.
We also highlighted the substantial performance improvements our tools offer,
as this is crucial for construction and simulation of models of whole cells and organisms,
which necessitate the simulation of significantly large-scale models.

@@

For the reaction networks,
our focus is on first-order differential equations
that describe the rate of change of species.
SimBio simplifies the definition of these network models by introducing `Species`,
which consists of a `poincare.Variable` and stoichiometry,
and `Reaction`,
a construct that converts reactants into products taking into account the stoichiometry (@fig-simbio).
Additionally, SimBio features `MassAction`, a subclass of `Reaction`,
which intuitively incorporates reactants into the rate law (@fig-simbio).
**TODO: is `RateLaw` not better term than `MassAction`?**

@@

**TODO: I amnto sure whether the number of figures are already over the limit, you may have to combine figures, while doing so, if space permits create a small schematics / summary figure, with icons and texts showing the significance / and applicability of the tools developed.**
